\section{Conclusion}\label{sec:conclusion}

In conclusion, we have simulated a 2D Ising model with no external magnetic field, by performing Monte Carlo simulation with the Metropolis algorithm. The model reproduces the expected existence of phase transitions near a critical temperature. We estimate the critical temperatures for four different lattice sizes, and use these values to obtain an estimate of the analytical value of the critical temperature for the 2D Ising model of infinite size. Our approximation reads $T_C(L=\infty) \approx 2.2651\tunit$, which compared to the analytical expression $T_C(L=\infty)\approx 2.2692\tunit$ is a fairly decent result, but slightly too low. With an improved algorithm better suited for critical temperatures (see section \ref{subsec_discussion:limitations}), and/or a more thorough analysis, as discussed in sections \ref{subsec_discussion:sample_size_equi} and \ref{subsec_discussion:critical_temps_curvefit}, the approximation of critical temperatures, and the efficiency of which, would probably be improved greatly. Our results do however lack a proper treatment of uncertainties. Although errors and model validity are discussed, without a proper treatment of the uncertainties we may only conclude that our obtained results are comparable estimates to the true analytical value.




\section{Acknowledgements}
We would like to offer our most sincere gratitude towards the tarmac in Blinderveien for its high bulk modulus, and the fragility of the human body. This work could not have been done without their combined support.   

% ===========================================